\chapter{Flight Dynamics and Control}
\label{chapter:papertwo}

\section{Coordinate Frames and System Definition}
\label{sec:coord_frames}

This chapter presents the flight dynamics and control framework for the
vehicle.  The critical design feature of this quadrotor is that
\emph{all four propellers spin in the same direction}.
The individual aerodynamic yaw torques therefore do not cancel, and the
vehicle sustains a non-zero net yaw torque during flight.  At equilibrium
this drives the vehicle to a constant spinning yaw rate---the
\emph{relaxed hovering condition}~\cite{mueller2016relaxed,bai2020splitflyer}---which
fundamentally shapes the dynamics and motivates the gyroscopic analysis that
follows.

\paragraph{Inertial and body frames.}
The right-handed inertial frame $(\hat{x}^w, \hat{y}^w, \hat{z}^w)$ is
fixed to the ground with $\hat{z}^w$ pointing vertically upward.  The
position of the vehicle's centre of mass (COM) in this frame is
$\boldsymbol{p} = [x,\,y,\,z]^{\top}$.  The right-handed body frame
$(\hat{x}^b, \hat{y}^b, \hat{z}^b)$ is rigidly attached to the vehicle with
its origin at the COM; the $\hat{z}^b$ axis is aligned with the common
thrust axis of the four propellers.  The rotation matrix
$\boldsymbol{R}\in SO(3)$ maps body-frame vectors to the inertial frame,
so the world-frame direction of the thrust axis is $\boldsymbol{R}\hat{e}_3$,
where $\hat{e}_3 = [0,0,1]^{\top}$.  The body angular velocity expressed in
the body frame is $\boldsymbol{\omega} = [\omega_x, \omega_y, \omega_z]^{\top}$.
The yaw angle $\psi$ is defined as the rotation of $\hat{x}^b$ about
$\hat{z}^w$ relative to $\hat{x}^w$.

\paragraph{Non-revolving frame.}
Because the vehicle spins at high yaw rate, a non-revolving intermediate
frame $(\hat{x}^m, \hat{y}^m, \hat{z}^b)$ is introduced to describe the
attitude of the spinning disk independently of the instantaneous yaw
angle~\cite{bai2022splitflyerair}.  Under the small-inclination assumption
$|\alpha_x|, |\alpha_y| \ll 1$, the reduced attitude state
$\boldsymbol{\eta} = [\alpha_x, \alpha_y]^{\top}$ parametrises the deviation
of $\hat{z}^b$ from vertical, and the non-revolving axes are approximated as
\begin{equation}
    \hat{x}^m \approx [1,\;0,\;{-\alpha_y}]^{\top},
    \qquad
    \hat{y}^m \approx [0,\;1,\;\alpha_x]^{\top}.
    \label{eq:nm_frame}
\end{equation}
The pair $(\hat{x}^m, \hat{y}^m)$ spans the horizontal plane for any yaw
angle $\psi$; this frame is the natural setting for the gyroscopic reduced
attitude dynamics derived in Section~\ref{sec:reduced_dynamics}.

\paragraph{Propeller geometry.}
The four propellers are arranged in an X-configuration at equal arm length
$l$ from the COM.  Following the convention of
\textcite{bai2022splitflyerair}, their body-frame position vectors are
\begin{equation}
    \boldsymbol{l}_1 = -\boldsymbol{l}_3
        = \frac{l}{\sqrt{2}}\!\left(\hat{x}^b + \hat{y}^b\right),
    \qquad
    \boldsymbol{l}_2 = -\boldsymbol{l}_4
        = \frac{l}{\sqrt{2}}\!\left(\hat{x}^b - \hat{y}^b\right).
    \label{eq:prop_locs}
\end{equation}
Each propeller $i$ produces a propulsive thrust $f_{i,p} \geq 0$ directed
along $+\hat{z}^b$ and an aerodynamic reaction torque $c\,f_{i,p}$ about
$\hat{z}^b$, where $c > 0$ is the common thrust-to-drag ratio (identical
for all four rotors since they share the same spinning direction).
Consequently the total yaw torque is
\begin{equation}
    \tau_z = c\sum_{i=1}^{4} f_{i,p} = c\,F,
    \label{eq:yaw_torque}
\end{equation}
where $F = \sum_i f_{i,p}$ is the total thrust.  Equation~\eqref{eq:yaw_torque}
shows that the yaw torque is always coupled to the total thrust and cannot
be nullified independently.  This coupling drives the vehicle passively to a
constant spin rate---the \emph{relaxed hovering condition}~\cite{mueller2016relaxed}---whose
equilibrium value is set by the balance between propeller reaction torque and
the drag of the spin-limiting plates, as derived in
Section~\ref{sec:attitude_dynamics}.

\paragraph{Inertia and drag.}
By design, mass is distributed symmetrically about the $\hat{z}^b$ axis so
that the moments of inertia about roll and pitch are approximately equal,
$I_x \approx I_y \equiv I_d$.  The full inertia tensor is therefore
\begin{equation}
    \boldsymbol{I} = \mathrm{diag}(I_d,\;I_d,\;I_z),
    \label{eq:inertia}
\end{equation}
which is the axisymmetric condition required to treat the vehicle as a
gyroscope in Section~\ref{sec:reduced_dynamics}.  Rotor-induced drag is
characterised by the coefficient matrix
$\boldsymbol{B} = \mathrm{diag}(B_h, B_h, B_v)$, and aerodynamic body
damping by $\boldsymbol{D} = \mathrm{diag}(D_h, D_h, D_v)$, where subscripts
$h$ and $v$ denote horizontal and vertical components
respectively~\cite{bai2022splitflyerair}.  The vehicle mass is $m = 44\;\mathrm{g}$;
remaining hardware parameters are specified in Chapter~\ref{chapter:paperone}.

\paragraph{Propeller aerodynamic model.}
Each propeller $i$ spins about its own thrust axis at blade speed $\Omega_i$,
which is distinct from and substantially larger than the vehicle yaw rate
$\omega_z$.  The propulsive thrust follows the standard aerodynamic model
\begin{equation}
    f_{i,p} = k_f\,\Omega_i^2,
    \label{eq:thrust_model}
\end{equation}
where $k_f > 0$ is a propeller constant.  Section~\ref{sec:attitude_dynamics}
introduces an additional aerodynamic effect absent from conventional
multirotor models: an advancing/retreating blade moment arising because the
four hydrofoils, mounted between the motor arms at $45^{\circ}$ azimuthal
offsets, sweep through the air as the entire vehicle yaws.  The hydrofoil
assembly acts as a single large rotor disk of radius $r_f$ rotating at the
vehicle yaw rate $\omega_z$; the analysis introduces the hydrofoil
advancing/retreating blade coefficient $K_{AB}$, parameterised by the
hydrofoil chord $c_f$, collective pitch $\theta_f$, two-dimensional lift
slope $a_0$, and foil tip radius $r_f$.

\section{Translational Dynamics}
\label{sec:translational_dynamics}

Treating the vehicle as a rigid body of mass $m$, Newton's second law gives
\begin{equation}
    m\ddot{\boldsymbol{p}} = \boldsymbol{R}\sum_{i=1}^{4}\boldsymbol{f}_i
        - mg\hat{e}_3,
    \label{eq:trans_eom}
\end{equation}
where the total force from propeller $i$ is
\begin{equation}
    \boldsymbol{f}_i = \hat{e}_3 f_{i,p} + \boldsymbol{f}_{i,d}.
    \label{eq:prop_force}
\end{equation}
The propulsive thrust $f_{i,p}$ is directed along the body $\hat{z}^b$ axis.
The rotor-induced drag $\boldsymbol{f}_{i,d}$, modelled as a linear function
of the air velocity at the propeller disk~\cite{bai2022splitflyerair}, is
\begin{equation}
    \boldsymbol{f}_{i,d} = -\boldsymbol{B}
        \!\left(\boldsymbol{R}^{-1}\dot{\boldsymbol{p}}
        + \boldsymbol{\omega}\times\boldsymbol{l}_i\right),
    \label{eq:rotor_drag}
\end{equation}
where the first term accounts for the translational velocity of the vehicle
and the second for the velocity of the propeller hub due to body rotation.

\paragraph{Cancellation of rotation-induced drag.}
The symmetry of the X-configuration---$\boldsymbol{l}_1 + \boldsymbol{l}_2
+ \boldsymbol{l}_3 + \boldsymbol{l}_4 = \mathbf{0}$, which follows
immediately from~\eqref{eq:prop_locs}---ensures that the rotation-induced
drag terms cancel when summed:
\begin{equation}
    \sum_{i=1}^{4}\left(\boldsymbol{\omega}\times\boldsymbol{l}_i\right)
        = \boldsymbol{\omega}\times\sum_{i=1}^{4}\boldsymbol{l}_i
        = \mathbf{0}.
    \label{eq:drag_cancel}
\end{equation}
This cancellation holds regardless of propeller spin direction and therefore
applies equally to our vehicle.  Substituting into~\eqref{eq:trans_eom}:
\begin{equation}
    m\ddot{\boldsymbol{p}}
        = \boldsymbol{R}\hat{e}_3 F
          - 4\boldsymbol{R}\boldsymbol{B}\boldsymbol{R}^{-1}\dot{\boldsymbol{p}}
          - mg\hat{e}_3.
    \label{eq:trans_full}
\end{equation}
The second term is a translational damping force present in all multirotor
vehicles.  For non-aggressive near-hover flight it is small relative to
gravity and is neglected in what follows~\cite{bai2022splitflyerair}, yielding
\begin{equation}
    m\ddot{\boldsymbol{p}} \approx \boldsymbol{R}\hat{e}_3 F - mg\hat{e}_3.
    \label{eq:trans_hover}
\end{equation}

\paragraph{Note on the advancing/retreating blade effect.}
The advancing/retreating blade moment derived in
Section~\ref{sec:attitude_dynamics} arises from the spinning hydrofoil
assembly generating an asymmetric vertical lift distribution when the vehicle
translates horizontally.  The first-harmonic lift asymmetry produces zero net
horizontal force by symmetry of the four-foil X-configuration
(the $\sum_i\sin\phi_i = \sum_i\cos\phi_i = 0$ identities cancel the
horizontal components), so the effect manifests solely as a torque about the
vehicle COM and does not appear in~\eqref{eq:trans_eom}.  The translational
dynamics~\eqref{eq:trans_hover} are therefore unaffected.

\paragraph{Near-hover linearisation.}
For small inclinations $|\alpha_x|,|\alpha_y|\ll 1$, the body $\hat{z}^b$
axis expressed in the inertial frame is approximately
\begin{equation}
    \boldsymbol{R}\hat{e}_3 \approx
        \begin{bmatrix} J\boldsymbol{\eta} \\ 1 \end{bmatrix},
    \qquad
    J = \begin{bmatrix} 0 & 1 \\ -1 & 0 \end{bmatrix},
    \label{eq:zb_approx}
\end{equation}
where $J$ is the $-90^{\circ}$ rotation matrix in the horizontal plane.
Splitting~\eqref{eq:trans_hover} into horizontal and vertical components
and using the hover thrust condition $F\approx mg$ for the horizontal term
gives
\begin{align}
    m\ddot{\boldsymbol{p}}_2 &= mg\,J\boldsymbol{\eta},
    \label{eq:horiz_dyn}\\
    m\ddot{z} &= F - mg,
    \label{eq:vert_dyn}
\end{align}
where $\boldsymbol{p}_2 = [x,\,y]^{\top}$ is the horizontal position.
Equation~\eqref{eq:horiz_dyn} shows that horizontal acceleration is driven
solely by the vehicle tilt $\boldsymbol{\eta}$, and forms the basis of the
position controller in Section~\ref{sec:controller}.

\section{Attitude Dynamics}
\label{sec:attitude_dynamics}

The rotational motion is governed by Euler's equations in the body frame.
The angular momentum equation is
\begin{equation}
    \boldsymbol{I}\dot{\boldsymbol{\omega}}
        + \boldsymbol{\omega}\times\boldsymbol{I}\boldsymbol{\omega}
        = \sum_{i=1}^{4}\boldsymbol{\tau}_i
          + \boldsymbol{\tau}_s,
    \label{eq:euler}
\end{equation}
where $\boldsymbol{\tau}_i$ is the net torque produced by propeller $i$
at the centre of mass, and $\boldsymbol{\tau}_s = -\boldsymbol{D}\boldsymbol{\omega}$
is the aerodynamic body damping.

\paragraph{Propeller torques.}
The total torque from propeller $i$ decomposes into a thrust-induced
component and a drag-induced component:
\begin{equation}
    \boldsymbol{\tau}_i = \boldsymbol{\tau}_{i,p} + \boldsymbol{\tau}_{i,d}.
    \label{eq:prop_torque_decomp}
\end{equation}
The thrust-induced torque combines the moment of the propulsive force about
the centre of mass with the aerodynamic reaction torque of the spinning
blade~\cite{bai2022splitflyerair}:
\begin{equation}
    \boldsymbol{\tau}_{i,p}
        = \boldsymbol{l}_i \times f_{i,p}\hat{e}_3
          + c\,f_{i,p}\hat{e}_3.
    \label{eq:tau_thrust}
\end{equation}
The first term generates the differential roll/pitch torques used for
attitude control; the second is the yaw reaction torque.  Summing the yaw
contributions across all four propellers recovers~\eqref{eq:yaw_torque},
showing that yaw remains coupled to the total thrust and is not independently
actuated.

The drag-induced torque arises from the rotor drag force
$\boldsymbol{f}_{i,d}$ (defined in~\eqref{eq:rotor_drag}) acting at the
moment arm $\boldsymbol{l}_i$:
\begin{equation}
    \boldsymbol{\tau}_{i,d}
        = \boldsymbol{l}_i \times \boldsymbol{f}_{i,d}
        = -\boldsymbol{l}_i
            \times \boldsymbol{B}
            \!\left(\boldsymbol{R}^{-1}\dot{\boldsymbol{p}}
                  + \boldsymbol{\omega}\times\boldsymbol{l}_i\right).
    \label{eq:tau_drag}
\end{equation}

\paragraph{Advancing/retreating blade moment.}
The four hydrofoils are mounted between the motor arms at $45^{\circ}$
azimuthal offsets, at azimuths $\phi_i \in \{0^{\circ}, 90^{\circ},
180^{\circ}, 270^{\circ}\}$ and equal tip radius $r_f$ from the vehicle COM
(distinct from the motor arm length $l$).  As the vehicle spins at yaw rate
$\omega_z$, each hydrofoil sweeps tangentially at speed $\omega_z r_f$.
The four-foil assembly therefore functions aerodynamically as a single large
rotor disk of radius $r_f$ rotating at $\omega_z$---a vehicle-scale rotor
distinct from the small propellers.

When the vehicle translates horizontally, this large rotor experiences the
classical advancing/retreating blade asymmetry.  Let
$\boldsymbol{v}_b = (\boldsymbol{R}^{-1}\dot{\boldsymbol{p}})_\perp$ denote
the horizontal translational velocity expressed in the body frame.  At foil
$i$, located at azimuth $\phi_i$ from $+\hat{x}^b$ and tangential direction
$\hat{\boldsymbol{t}}_i = (-\sin\phi_i,\,\cos\phi_i,\,0)^{\top}$, the total
tangential airspeed is
\begin{equation}
    U_i = \omega_z r_f + \hat{\boldsymbol{t}}_i\cdot\boldsymbol{v}_b.
    \label{eq:foil_speed}
\end{equation}
The first term $\omega_z r_f$ is the primary sweep speed; the second is the
translational perturbation that breaks the rotational symmetry.  Modelling
each hydrofoil as a flat plate with chord $c_f$, collective pitch $\theta_f$,
and two-dimensional lift slope $a_0$, the aerodynamic lift in air is
\begin{equation}
    L_i = k_{HF}\,U_i^2,
    \qquad
    k_{HF} = \tfrac{1}{2}\rho\,c_f\,a_0\,\theta_f.
    \label{eq:foil_lift}
\end{equation}
The moment about the COM from the vertical lift force $L_i\hat{e}_3$ at
foil position $r_f(\cos\phi_i,\,\sin\phi_i,\,0)^{\top}$ is
\begin{equation}
    \boldsymbol{\tau}_i = L_i\,r_f
        \bigl(\sin\phi_i\,\hat{x}^b - \cos\phi_i\,\hat{y}^b\bigr).
    \label{eq:foil_moment_i}
\end{equation}
Expanding $U_i^2 \approx (\omega_z r_f)^2 + 2\omega_z r_f\,(\hat{\boldsymbol{t}}_i\cdot\boldsymbol{v}_b)$
to first order in $|\boldsymbol{v}_b|/(\omega_z r_f)$ and summing over the
four foils at $\phi_i \in \{0^{\circ}, 90^{\circ}, 180^{\circ}, 270^{\circ}\}$,
the zeroth-order terms cancel by the identities
$\sum_i\sin\phi_i = \sum_i\cos\phi_i = 0$, and the first-order cross-terms
combine using $\sum_i\sin^2\phi_i = \sum_i\cos^2\phi_i = 2$ and
$\sum_i\sin\phi_i\cos\phi_i = 0$.  The result is
\begin{equation}
    \boldsymbol{\tau}_{AB}
        = -K_{AB}
            \bigl(\boldsymbol{R}^{-1}\dot{\boldsymbol{p}}\bigr)_\perp,
    \label{eq:tau_ab}
\end{equation}
where the \emph{hydrofoil advancing/retreating blade coefficient} is
\begin{equation}
    K_{AB} = 4\,k_{HF}\,\omega_z\,r_f^2
           = 2\rho\,c_f\,a_0\,\theta_f\,\omega_z\,r_f^2.
    \label{eq:KAB}
\end{equation}
The coefficient $K_{AB}$ scales with $\omega_z$: the faster the vehicle
spins (heavier vehicle or thinner air, from~\eqref{eq:omegaz_hover}), the
stronger the translation--attitude coupling.  The negative sign
in~\eqref{eq:tau_ab} indicates that the moment opposes the translational
velocity in the body frame: when the vehicle moves in $+\hat{x}^b$, the
foil at $\phi_i = 270^{\circ}$ (advancing, on the $-\hat{y}^b$ side for
$c > 0$) generates more lift, rolling the vehicle toward $-\hat{x}^b$.
This translation--attitude coupling is absent from conventional multirotor
models and propagates into the reduced attitude dynamics of
Section~\ref{sec:reduced_dynamics}.

\paragraph{Complete Euler equation.}
Collecting all torque contributions, with the propeller torques
$\boldsymbol{\tau}_{i,p}$ and $\boldsymbol{\tau}_{i,d}$ acting at each
rotor hub and the hydrofoil moment $\boldsymbol{\tau}_{AB}$ acting on the
vehicle as a whole:
\begin{equation}
    \boldsymbol{I}\dot{\boldsymbol{\omega}}
        + \boldsymbol{\omega}\times\boldsymbol{I}\boldsymbol{\omega}
        = \sum_{i=1}^{4}
            \bigl(\boldsymbol{\tau}_{i,p}
                  + \boldsymbol{\tau}_{i,d}\bigr)
          + \boldsymbol{\tau}_{AB}
          + \boldsymbol{\tau}_s.
    \label{eq:euler_full}
\end{equation}

\paragraph{Passive yaw equilibrium.}
The hydrofoil advancing/retreating blade torque $\boldsymbol{\tau}_{AB}$
lies entirely in the body $\hat{x}^b$--$\hat{y}^b$ plane and carries no
$\hat{z}^b$ component.  The dominant source of yaw drag is instead a set of
four flat \emph{drag plates}, mounted vertically with their faces parallel to
the yaw axis so that they sweep broadside through the air as the vehicle
spins.  At yaw rate $\omega_z$, the plate on each arm, with area centroid at
effective radius $r_d$, moves at tangential speed $\omega_z r_d$, and its
drag, acting opposite to the sweep direction, is
\begin{equation}
    D_{d,i} = k_{D,d}\,(\omega_z r_d)^2,
    \qquad
    k_{D,d} = \tfrac{1}{2}\rho\,C_{D,d}\,S_d,
    \label{eq:plate_drag}
\end{equation}
where $C_{D,d}$ is the plate drag coefficient and $S_d$ its planform area.
For a vehicle operating at a fixed altitude, $\rho$ is a constant of the
environment; $k_{D,d}$ is therefore determined entirely by the plate geometry
($C_{D,d}$, $S_d$) and is a fixed design parameter.  Each drag force acts at
moment arm $r_d$ about the $\hat{z}^b$ axis, contributing a yaw-retarding
torque $-k_{D,d}r_d^3\omega_z^2$ per plate.  Extracting the $\hat{z}^b$
projection of~\eqref{eq:euler_full}, using
$[\boldsymbol{\omega}\times\boldsymbol{I}\boldsymbol{\omega}]_z = 0$
(axisymmetric inertia), and collecting all four plate drag torques gives the
yaw dynamics
\begin{equation}
    I_z\dot{\omega}_z
        = c\,F
          - 4\,k_{D,d}\,r_d^3\,\omega_z^2.
    \label{eq:yaw_ode}
\end{equation}
The quadratic drag on the right-hand side is set by the plate geometry; it
creates a unique stable equilibrium at $\omega_z^* > 0$.  At sustained hover,
$F = mg$, giving the \emph{relaxed hover spin rate}
\begin{equation}
    \boxed{\omega_z^* = \sqrt{\frac{c\,m\,g}{4\,k_{D,d}\,r_d^3}}.}
    \label{eq:omegaz_hover}
\end{equation}
This equilibrium is reached entirely passively: the plate drag automatically
balances the propeller yaw torque without any active control input.  The
plates are sized so that $\omega_z^*$ stays below the controllability ceiling
of $20\pi\;\mathrm{rad\,s^{-1}}$ imposed by the $100\;\mathrm{Hz}$ control
loop (Section~\ref{sec:design_requirements}); for the present vehicle the
resulting spin rate is $\omega_z^* \approx 52\;\mathrm{rad\,s^{-1}}$
($3000^{\circ}\,\mathrm{s^{-1}}$).  The hydrofoils contribute a smaller
profile-drag term of the same quadratic form, which can be absorbed into
$k_{D,d}$ without altering the structure of the balance.

\paragraph{Geometric determinants of the spin rate.}
Equation~\eqref{eq:omegaz_hover} expresses $\omega_z^*$ as the balance
between two geometric quantities: the propeller reaction coefficient $c$
(yaw drive per unit thrust) and the plate drag stiffness $4k_{D,d}r_d^3$
(drag torque per unit $\omega_z^2$).  For a given vehicle at a given
altitude both are fixed design parameters, so $\omega_z^*$ is a direct
consequence of the geometry---not of any control input.

The plate radius $r_d$ is particularly influential: the drag torque scales as
$r_d^3$ (moment arm $r_d$ multiplied by drag force $\propto (\omega_z r_d)^2$),
so placing the plates further from the centre slows the spin strongly.
Increasing the plate area $S_d$ or its drag coefficient $C_{D,d}$ has the same
slowing effect through $k_{D,d}$.  These plate parameters are the primary
levers for tuning $\omega_z^*$ to sit below the controllability ceiling.
Conversely, increasing the propeller reaction-to-thrust ratio $c$ speeds the
vehicle up.

Vehicle weight enters through the hover thrust: a heavier vehicle requires
more thrust and therefore generates more yaw drive torque $cF = cmg$,
spinning \emph{faster}.  The square-root relationship makes this sensitivity
mild---doubling the mass raises $\omega_z^*$ by only $\sqrt{2}$---and means
the spin rate is robust to small mass variations.

This constant $\omega_z^*$ is the spin rate assumed throughout
Section~\ref{sec:reduced_dynamics}.

\section{Reduced Dynamics under High Yaw Rate}
\label{sec:reduced_dynamics}

The full Euler equation~\eqref{eq:euler_full} governs all three rotational
degrees of freedom simultaneously.  The equilibrium analysis of
Section~\ref{sec:attitude_dynamics} established that the vehicle operates at
a constant yaw rate $\omega_z^*$ satisfying~\eqref{eq:omegaz_hover}, with
the tilt rates $|\omega_x|, |\omega_y|$ remaining small compared to
$\omega_z^*$.  This separation of timescales motivates a reduced
two-dimensional description in the tilt state $\boldsymbol{\eta}$ that
eliminates the fast yaw and retains only the controllable attitude evolution.

\paragraph{Inertial angular momentum.}
The angular momentum in the inertial frame is
\begin{equation}
    \mathbf{h} = \boldsymbol{R}\boldsymbol{I}\boldsymbol{\omega}
        = I_d\boldsymbol{R}[\omega_x,\;\omega_y,\;0]^{\top}
          + I_z\omega_z^*\boldsymbol{R}\hat{e}_3.
    \label{eq:inertial_h}
\end{equation}
The second term is the large gyroscopic spin momentum $I_z\omega_z^*$
directed along the tilting thrust axis $\boldsymbol{R}\hat{e}_3$.
By Newton's second law in the inertial frame,
\begin{equation}
    \dot{\mathbf{h}}
        = \boldsymbol{R}\!\left(
            \sum_{i=1}^{4}\bigl(\boldsymbol{\tau}_{i,p}
                               + \boldsymbol{\tau}_{i,d}\bigr)
            + \boldsymbol{\tau}_{AB}
            + \boldsymbol{\tau}_s
          \right).
    \label{eq:h_dot}
\end{equation}

\paragraph{Horizontal projection and gyroscopic coupling.}
Projecting~\eqref{eq:h_dot} onto the inertial horizontal plane via
$[e_1,e_2]^{\top}$ and using $[\omega_x, \omega_y]^{\top} \approx
\dot{\boldsymbol{\eta}}$ for small tilt, the two parts
of~\eqref{eq:inertial_h} contribute separately.  The tilt momentum term
gives $I_d\ddot{\boldsymbol{\eta}}$ on differentiation.  The spin momentum
term $I_z\omega_z^*\boldsymbol{R}\hat{e}_3$ precesses as the body tilts:
\begin{equation}
    [e_1,e_2]^{\top}
    \frac{\mathrm{d}}{\mathrm{d}t}
    \bigl(I_z\omega_z^*\boldsymbol{R}\hat{e}_3\bigr)
    = I_z\omega_z^*[e_1,e_2]^{\top}
      \boldsymbol{R}\bigl(\boldsymbol{\omega}\times\hat{e}_3\bigr)
    \approx I_z\omega_z^*\boldsymbol{J}\dot{\boldsymbol{\eta}},
    \label{eq:gyro_proj}
\end{equation}
where the last step uses $\boldsymbol{\omega}\times\hat{e}_3 \approx
\boldsymbol{J}\boldsymbol{\omega}_h = [\omega_y,-\omega_x]^{\top}$ and the
fact that $\boldsymbol{J}$ commutes with planar rotations (since $\boldsymbol{J}$
is itself a $-90^{\circ}$ rotation in the horizontal
plane)~\cite{bai2022splitflyerair}.  The result is the \emph{gyroscopic
coupling} term $I_z\omega_z^*\boldsymbol{J}\dot{\boldsymbol{\eta}}$: the
large spin angular momentum precesses in the inertial plane as the vehicle
tilts, coupling $\dot{\alpha}_x$ into the $\alpha_y$ equation and vice
versa.

\paragraph{Projected external torques.}
Three torque groups appear on the right side of~\eqref{eq:h_dot}.

\noindent\textit{Control torques.}
The inertial projection of the thrust-induced roll/pitch torques defines the
\emph{reduced control input}:
\begin{equation}
    \tilde{\boldsymbol{\tau}}_w
        = [e_1,e_2]^{\top}\boldsymbol{R}
          \sum_{i=1}^{4}\bigl(\boldsymbol{l}_i\times f_{i,p}\hat{e}_3\bigr).
    \label{eq:tau_w_reduced}
\end{equation}
The yaw reaction torques $cf_{i,p}\hat{e}_3$ project to $cF\boldsymbol{J}\boldsymbol{\eta}$,
which is second-order in tilt amplitude and is neglected.

\noindent\textit{Drag torques.}
The horizontal arm-sweep component of the rotor drag produces only yaw
torque (its roll/pitch contribution cancels by
symmetry~\eqref{eq:drag_cancel}).  The vertical rotor drag at each hub,
arising from the tilt-induced hub velocity $(\boldsymbol{\omega}\times
\boldsymbol{l}_i)_z$, contributes a roll/pitch damping; summing over four
arms of equal length gives $-2B_v l^2\dot{\boldsymbol{\eta}}$.  Adding the
body damping $-D_h\dot{\boldsymbol{\eta}}$ from $\boldsymbol{\tau}_s$ defines
the \emph{effective tilt damping coefficient}
\begin{equation}
    \bar{D} \triangleq 2B_v l^2 + D_h,
    \label{eq:Dbar}
\end{equation}
so the projected drag torque is $\tilde{\boldsymbol{\tau}}_d =
-\bar{D}\dot{\boldsymbol{\eta}}$.

\noindent\textit{Hydrofoil advancing/retreating blade torque.}
From~\eqref{eq:tau_ab}, the advancing/retreating blade moment in the body
frame is $\boldsymbol{\tau}_{AB} = -K_{AB}(\boldsymbol{R}^{-1}
\dot{\boldsymbol{p}})_\perp$.  Projecting to the inertial horizontal plane
near hover, where $\boldsymbol{R}(\boldsymbol{R}^{-1}\dot{\boldsymbol{p}})_\perp
\approx \dot{\boldsymbol{p}}_2$ for small tilt, gives
\begin{equation}
    \tilde{\boldsymbol{\tau}}_{AB}
        = [e_1,e_2]^{\top}\boldsymbol{R}\,\boldsymbol{\tau}_{AB}
        \approx -K_{AB}\dot{\boldsymbol{p}}_2.
    \label{eq:tau_AB_reduced}
\end{equation}

\paragraph{Reduced attitude equation.}
Combining the projected angular momentum with all external torques yields
the \emph{gyroscopic reduced attitude equation}:
\begin{equation}
    \boxed{I_d\ddot{\boldsymbol{\eta}}
        + I_z\omega_z^*\boldsymbol{J}\dot{\boldsymbol{\eta}}
        = \tilde{\boldsymbol{\tau}}_w
          - \bar{D}\dot{\boldsymbol{\eta}}
          - K_{AB}\dot{\boldsymbol{p}}_2.}
    \label{eq:reduced_attitude}
\end{equation}
The gyroscopic term $I_z\omega_z^*\boldsymbol{J}\dot{\boldsymbol{\eta}}$
couples the two tilt channels through precession of the spin angular momentum;
$-\bar{D}\dot{\boldsymbol{\eta}}$ is the projected aerodynamic damping;
and $-K_{AB}\dot{\boldsymbol{p}}_2$ is the hydrofoil advancing/retreating
blade term from Section~\ref{sec:attitude_dynamics}, which injects the
horizontal translational velocity as a disturbance into the tilt dynamics.

Together with the horizontal position equation~\eqref{eq:horiz_dyn}:
\begin{equation}
    m\ddot{\boldsymbol{p}}_2 = mg\boldsymbol{J}\boldsymbol{\eta},
    \label{eq:horiz_coupled}
\end{equation}
equations~\eqref{eq:reduced_attitude}--\eqref{eq:horiz_coupled} form a
coupled fourth-order system in $(\boldsymbol{p}_2,\boldsymbol{\eta})$ driven
by $\tilde{\boldsymbol{\tau}}_w$.  This is the plant model for the controller
developed in Section~\ref{sec:controller}.

\paragraph{Actuation: degrees of freedom and thrust generation.}
The reduced input $\tilde{\boldsymbol{\tau}}_w$ is a two-component vector
realised by the four propeller thrusts.  The four independently commandable
propellers provide three effective degrees of freedom: total thrust $F$,
body-frame roll torque $\tau_x^b$, and body-frame pitch torque $\tau_y^b$.
Any desired $\tilde{\boldsymbol{\tau}}_w$ can be generated \emph{statically}
by a fixed differential thrust pattern, with no time-varying modulation of
the thrust commands.  The explicit mapping is
\begin{equation}
    \begin{bmatrix}F\\\tau_x^b\\\tau_y^b\end{bmatrix}
    = \boldsymbol{M}
      \begin{bmatrix}f_{1,p}\\f_{2,p}\\f_{3,p}\\f_{4,p}\end{bmatrix},
    \qquad
    \boldsymbol{M} = \begin{bmatrix}
        1 & 1 & 1 & 1 \\
        \tfrac{l}{\sqrt{2}} & -\tfrac{l}{\sqrt{2}} & -\tfrac{l}{\sqrt{2}} & \tfrac{l}{\sqrt{2}} \\
        -\tfrac{l}{\sqrt{2}} & -\tfrac{l}{\sqrt{2}} & \tfrac{l}{\sqrt{2}} & \tfrac{l}{\sqrt{2}}
    \end{bmatrix},
    \label{eq:input_map}
\end{equation}
obtained from~\eqref{eq:tau_thrust} for the X-configuration~\eqref{eq:prop_locs}.
The one-dimensional null space of $\boldsymbol{M}$ can be exploited to
satisfy the positivity constraints $f_{i,p}\geq 0$ or to minimise total
power, as discussed in Section~\ref{sec:controller}.  Yaw remains passive
($\tau_z = cF$) throughout.

\section{Controller Design}
\label{sec:controller}

The control objective is to track a desired horizontal trajectory
$\boldsymbol{p}_{2,d}(t)$ and altitude set-point $z_d$ while maintaining the
passive relaxed-hover spin of Section~\ref{sec:attitude_dynamics}.  The yaw
degree of freedom requires no active control.

\paragraph{Non-cascaded horizontal controller.}
Define the horizontal position error $\tilde{\boldsymbol{e}} =
\boldsymbol{p}_2 - \boldsymbol{p}_{2,d}$.  From~\eqref{eq:horiz_coupled},
the attitude and its rate are related to position derivatives by
\begin{equation}
    \boldsymbol{\eta} = -\frac{1}{g}\boldsymbol{J}\ddot{\boldsymbol{p}}_2,
    \qquad
    \dot{\boldsymbol{\eta}} = -\frac{1}{g}\boldsymbol{J}\boldsymbol{p}_2^{(3)},
    \label{eq:eta_from_p2}
\end{equation}
so the system~\eqref{eq:reduced_attitude}--\eqref{eq:horiz_coupled} is fourth
order in $\boldsymbol{p}_2$.  Solving~\eqref{eq:reduced_attitude} for
$\tilde{\boldsymbol{\tau}}_w$ yields the plant input--output relation:
\begin{equation}
    \tilde{\boldsymbol{\tau}}_w
        = -\frac{I_d}{g}\boldsymbol{J}\boldsymbol{p}_2^{(4)}
          + \bar{D}\dot{\boldsymbol{\eta}}
          + I_z\omega_z^*\boldsymbol{J}\dot{\boldsymbol{\eta}}
          + K_{AB}\dot{\boldsymbol{p}}_2.
    \label{eq:ctrl_inversion}
\end{equation}
The last three terms are the known plant dynamics---aerodynamic damping,
gyroscopic coupling, and hydrofoil blade coupling---to be cancelled
by the control law.

A fourth-order polynomial constraint is imposed on the error:
\begin{equation}
    \tilde{\boldsymbol{e}}^{(4)}
        + \lambda_3\tilde{\boldsymbol{e}}^{(3)}
        + \lambda_2\ddot{\tilde{\boldsymbol{e}}}
        + \lambda_1\dot{\tilde{\boldsymbol{e}}}
        + \lambda_0\tilde{\boldsymbol{e}} = \boldsymbol{0},
    \label{eq:error_poly}
\end{equation}
which is asymptotically stable when $s^4 + \lambda_3 s^3 + \lambda_2 s^2
+ \lambda_1 s + \lambda_0$ is Hurwitz~\cite{bai2022splitflyerair}.  The
constraint prescribes the reference fourth derivative:
\begin{equation}
    \boldsymbol{p}_2^{(4)}\big|_r
        = \boldsymbol{p}_{2,d}^{(4)}
          - \lambda_3\tilde{\boldsymbol{e}}^{(3)}
          - \lambda_2\ddot{\tilde{\boldsymbol{e}}}
          - \lambda_1\dot{\tilde{\boldsymbol{e}}}
          - \lambda_0\tilde{\boldsymbol{e}}.
    \label{eq:p2_ref}
\end{equation}
Substituting~\eqref{eq:p2_ref} into~\eqref{eq:ctrl_inversion} gives the
horizontal control torque:
\begin{equation}
    \tilde{\boldsymbol{\tau}}_{w,r}
        = -\frac{I_d}{g}\boldsymbol{J}\boldsymbol{p}_2^{(4)}\big|_r
          + \bar{D}\dot{\boldsymbol{\eta}}
          + I_z\omega_z^*\boldsymbol{J}\dot{\boldsymbol{\eta}}
          + K_{AB}\dot{\boldsymbol{p}}_2.
    \label{eq:ctrl_law}
\end{equation}
The first term is the error-driven command from~\eqref{eq:p2_ref}; the
remaining terms cancel the known plant dynamics.  The jerk
$\tilde{\boldsymbol{e}}^{(3)}$ in~\eqref{eq:p2_ref} is evaluated as
$g\boldsymbol{J}\dot{\boldsymbol{\eta}} - \boldsymbol{p}_{2,d}^{(3)}$
from~\eqref{eq:eta_from_p2}, so all feedback signals reduce to
$(\boldsymbol{p}_2,\,\dot{\boldsymbol{p}}_2,\,\boldsymbol{\eta},\,
\dot{\boldsymbol{\eta}})$---quantities available from position sensing and
an IMU.

\paragraph{Altitude controller.}
The vertical dynamics~\eqref{eq:vert_dyn} are controlled independently by
the total thrust.  A proportional-derivative law with gravity compensation
gives
\begin{equation}
    F_r = m\bigl[
        \ddot{z}_d
        + k_{v,z}(\dot{z}_d - \dot{z})
        + k_{p,z}(z_d - z)
    \bigr] + mg,
    \label{eq:alt_ctrl}
\end{equation}
where $k_{p,z},\,k_{v,z} > 0$ are the altitude proportional and derivative
gains.

\paragraph{Input allocation.}
The reference torque $\tilde{\boldsymbol{\tau}}_{w,r}$ is expressed in the
inertial horizontal frame; its body-frame equivalent is
\begin{equation}
    \begin{bmatrix}\tau_x^b\\\tau_y^b\end{bmatrix}
        = \boldsymbol{R}_{z,2}^{\top}(\psi)\,\tilde{\boldsymbol{\tau}}_{w,r},
    \label{eq:torque_projection}
\end{equation}
where $\boldsymbol{R}_{z,2}(\psi)$ is the $2\times 2$ yaw-rotation submatrix
of $\boldsymbol{R}$.  The individual propeller thrust commands are then
obtained from the minimum-norm solution of the mapping~\eqref{eq:input_map}:
\begin{equation}
    \boldsymbol{f}_r
        = \boldsymbol{M}^{\dagger}
          \begin{bmatrix}F_r\\\tau_x^b\\\tau_y^b\end{bmatrix},
    \qquad
    \boldsymbol{M}^{\dagger}
        = \boldsymbol{M}^{\top}
          \!\left(\boldsymbol{M}\boldsymbol{M}^{\top}\right)^{-1}.
    \label{eq:alloc}
\end{equation}
If any $f_{i,r} < 0$, the null vector
$\boldsymbol{n} \propto [1,\,-1,\,1,\,-1]^{\top}$ of $\boldsymbol{M}$
provides a one-parameter family of solutions: adding any scalar multiple of
$\boldsymbol{n}$ to $\boldsymbol{f}_r$ shifts all four thrusts equally
without altering $(F_r,\,\tau_x^b,\,\tau_y^b)$, and can always be chosen
to satisfy $f_{i,r} \geq 0$.