\chapter{Introduction}
\label{chapter:introduction}

\section{Background and Motivation}
\label{sec:background_motivation}

Micro aerial vehicles (MAVs) have become capable platforms for environmental
monitoring, search and rescue, and inspection, yet the overwhelming majority
operate exclusively in the air.  Many of the environments where small robots
are most useful---rivers, lakes, flooded terrain, and coastal zones---are
dominated by water, and a vehicle confined to flight must expend continuous
power to remain aloft above them.  Extending an MAV's mobility to include the
water surface would broaden its operational envelope to tasks such as water
sampling, surface skimming, and persistent monitoring of aquatic environments,
while allowing the vehicle to rest on or interact with the surface between
flight phases.% TODO: cite MAV applications / amphibious aerial robots

The central difficulty is not resting on water but leaving it again.  For a
small, weight- and power-constrained vehicle, repeatedly taking off from the
water surface is energetically demanding: overcoming surface adhesion and
accelerating from rest into flight typically requires either oversized thrust
or dedicated launch mechanisms that add mass and complexity.  As a result,
most amphibious aerial robots treat water contact as a terminal
state---landing, floating, or sampling---rather than as a mode of locomotion
in its own right.% TODO: cite water-takeoff mechanisms / aquatic-aerial robots

A different possibility is suggested by stone-skipping.  A flat stone thrown
at a shallow angle skips across water because, during each brief contact, the
water produces a large transient hydrodynamic reaction that redirects the
stone upward before it can sink; the contact is short, the energy lost per
bounce is small, and the stone traverses the surface in a sequence of low-cost
hops.% TODO: cite stone-skipping physics (e.g. Bocquet 2003)
This thesis adapts that principle to a powered vehicle: rather than hovering
above the water or floating on it, the vehicle hops across the surface, using
brief hydrofoil contact to recover much of the energy needed to return to the
air.  The mechanism that makes this possible is described in
Section~\ref{sec:water_skipping_mechanism}.

\section{Water-Skipping Mechanism}
\label{sec:water_skipping_mechanism}

The vehicle is a quadrotor whose four rotors all spin in the same direction.
Because their aerodynamic reaction torques do not cancel, the airframe itself
rotates continuously about its vertical axis during flight, settling at a
constant spin rate.  Mounted beneath the body, one under each arm, are four
hydrofoils that rotate rigidly with the airframe; as the vehicle spins, each
hydrofoil sweeps through a circular path at a tangential speed set by the spin
rate and its radius from the centre.

A single hop proceeds in three phases.  The vehicle first climbs and then
enters a brief free-fall, descending toward a still water surface.  At contact
the spinning hydrofoils slice into the water, and the combination of their
tangential sweep speed and their inclined attack angle generates a large
hydrodynamic lift.  Because the foils sweep continuously and the submersion
lasts only a few milliseconds, the resulting vertical impulse arrests the
descent and ejects the vehicle back into the air, where it recovers altitude
and repeats the cycle.

What makes this efficient is the contrast between the two media the hydrofoils
move through.  Water is nearly three orders of magnitude denser than air, so
the same hydrofoil, sweeping at the same speed, produces a far larger lift
force during its brief submersion than it does while spinning in air.  The
foils therefore impose almost no penalty on flight---they generate negligible
aerodynamic lift as the vehicle hovers and spins---yet deliver a powerful
upward impulse the instant they touch water.  This asymmetry allows each hop
to recover a large fraction of the mechanical energy that would otherwise be
spent climbing back out of the water, so the vehicle can sustain repeated hops
on modest power rather than paying the full cost of a water takeoff at every
bounce.

The vehicle's continuous spin thus serves a dual purpose: it places the
airframe in the relaxed-hover flight regime analysed in
Chapter~\ref{chapter:papertwo}, and it supplies the tangential blade velocity
that the hydrofoils require to generate lift on water contact.  The quantitative
model of the submersion and ejection is developed in Chapter~\ref{dynamic}, and
the physical realisation of the spinning airframe and its hydrofoils is
presented in Chapter~\ref{chapter:paperone}.

\section{Related Work}
\label{sec:related_work}

\paragraph{Aquatic--aerial robots.}
Robots able to operate in both air and water have been pursued from several
directions.  One line of work develops fixed-wing vehicles that plunge into
the water and relaunch from it~\cite{??}, while another adapts rotorcraft to
land on, float on, or dive beneath the surface~\cite{??}.  A third,
bio-inspired line mimics animals such as ?? that transition between swimming
and flight~\cite{??}.  Across these efforts, water contact is generally
treated as a destination---a dive, a landing, or a sampling stop---and the
return to the air is handled either by fully restarting flight or by a
dedicated launch event.

\paragraph{Leaving the water.}
Departing the surface is the most demanding part of amphibious operation, and
existing vehicles address it in energy-intensive ways.  Reported mechanisms
include impulsive jumping~\cite{??}, rapid gas generation or
combustion~\cite{??}, and simply oversizing the propulsion system to overcome
surface adhesion and accelerate from rest~\cite{??}.  These approaches enable a
single transition but are poorly suited to repeated water contact, because
each takeoff consumes a large, non-recoverable share of the energy budget.

\paragraph{Water-entry and skipping physics.}
The physics of objects bouncing on water has been studied independently of
robotics.  Stone-skipping and the related water-entry problem have been
analysed for both spinning and non-spinning bodies~\cite{??}, establishing
that a flat object striking water at a shallow angle experiences a large,
short-lived hydrodynamic reaction that can redirect it upward before it
sinks~\cite{??}.  These studies characterise the conditions for a successful
skip---approach angle, speed, and spin---but consider passive, unpowered
projectiles rather than a controlled, self-propelled vehicle.

\paragraph{Spinning multirotors.}
Multirotors that hover while spinning rapidly about their vertical axis have
been studied as relaxed-hover platforms.  Mueller and D'Andrea characterise
relaxed-hover solutions in which a multicopter sustains a constant spin rather
than holding a fixed heading~\cite{mueller2016relaxed}, and the SplitFlyer
vehicles realise a rotating multirotor whose attitude is governed by
gyroscopic reduced-attitude dynamics~\cite{bai2020splitflyer,bai2022splitflyerair}.
The flight control developed in this thesis operates in the same fast-spin
regime; its role here is to keep the spinning airframe stable and controllable
so that it can perform the water-skipping hop.

\paragraph{Summary.}
These bodies of work supply the individual ingredients---amphibious airframes,
water-entry hydrodynamics, and control of fast-spinning multirotors---but none
combines them into a vehicle that uses spinning hydrofoils to hop repeatedly
across water while recovering much of the energy of each bounce.  That
combination is the subject of this thesis.

\section{Research Objectives}

\section{Thesis Contributions}

\section{Thesis Organization}

The remainder of this thesis is organized as follows:

\textbf{Chapter~\ref{chapter:background}} provides essential background information on [research domain], including historical context, current state-of-the-art, and key challenges that motivate this work.

\textbf{Chapter~\ref{chapter:paperone}} presents our first major contribution: [brief description of paper one]. This chapter introduces [key concept 1] and demonstrates its effectiveness through comprehensive evaluation.

\textbf{Chapter~\ref{chapter:papertwo}} describes our second major contribution: [brief description of paper two]. This work builds upon the foundation established in Chapter~\ref{chapter:paperone} and addresses [specific problem area].

\textbf{Chapter~\ref{chapter:paperthree}} presents our third major contribution: [brief description of paper three]. This chapter integrates the previous contributions into a unified framework and demonstrates its effectiveness in real-world scenarios.

\textbf{Chapter~\ref{chapter:relatedwork}} provides a comprehensive survey of related work in [research domain], positioning our contributions within the broader research landscape.

\textbf{Chapter~\ref{chapter:conclusion}} summarizes the key contributions of this thesis, discusses their implications, and outlines promising directions for future research.
