\begin{publications}
\begin{enumerate}

\item \textbf{Your Name}, Co-Author A, Co-Author B, and Senior Author. Title of Your First Paper: A Novel Approach to [Research Problem]. In \textit{Proceedings of the International Conference on [Conference Name] ([Conference Abbreviation])}, [Month] [Day-Day], [Year].

\item Co-Author A*, \textbf{Your Name}*, Co-Author B, and Senior Author. Title of Your Second Paper: Advanced Techniques for [Research Area]. In \textit{IEEE Transactions on [Journal Name]}, vol. [Volume], no. [Number], pp. [Page Range], [Year].

* \textit{Both authors contributed equally to this work.}

\item \textbf{Your Name}, Co-Author C, Co-Author D, Senior Author A, and Senior Author B. Title of Your Third Paper: Scalable Solutions for [Application Domain]. In \textit{Proceedings of the [Conference Name] ([Conference Abbreviation])}, [Month] [Day-Day], [Year].

\item Co-Author E, \textbf{Your Name}, Co-Author F, and Senior Author. Title of Your Fourth Paper: Optimization Strategies in [Research Domain]. In \textit{[Journal Name]}, vol. [Volume], no. [Number], pp. [Page Range], [Year].

\item \textbf{Your Name} and Senior Author. Title of Your Fifth Paper: A Comprehensive Study of [Research Topic]. \textit{Under review at [Journal/Conference Name]}, [Year].

\item Co-Author G, Co-Author H, \textbf{Your Name}, and Senior Author. Title of Your Sixth Paper: Practical Implementation of [System Name]. In \textit{Proceedings of the [Workshop Name] at [Main Conference]}, [Month] [Year].

\end{enumerate}

\end{publications}