\chapter{Design of the Water-Skipping MAV}
\label{chapter:paperone}

\section{Design Requirements}
\label{sec:design_requirements}

The vehicle must satisfy two coupled sets of requirements: it must generate
enough lift to hover as a quadrotor, and its spinning hydrofoils must produce
sufficient hydrodynamic impulse during a brief water-surface contact to eject
it back into the air.

\paragraph{Flight requirements.}
The target all-up mass is $m = 44\;\mathrm{g}$.  All four propellers spin in
the same direction, so yaw torque cannot be independently cancelled; the
vehicle instead reaches the relaxed-hover equilibrium spin rate
$\omega_z^*$ derived in Chapter~\ref{chapter:papertwo}.  The attitude
controller runs at a loop rate of $100\;\mathrm{Hz}$, and to remain
controllable the spin must be slow enough to leave several control updates
within each revolution; the spin rate is therefore constrained to
$\omega_z^* < 20\pi\;\mathrm{rad\,s^{-1}}$ ($\approx 62.8\;\mathrm{rad\,s^{-1}}$,
i.e.\ at least ten control updates per revolution; equation~\eqref{eq:omegaz_hover}).
Because the same-direction propellers would otherwise drive the vehicle well
above this limit, dedicated drag plates (Section~\ref{sec:foil_design}) are
added to raise the passive yaw drag and hold $\omega_z^*$ within the
controllable range.  In operation the spin settles near
$\omega_z^* \approx 52\;\mathrm{rad\,s^{-1}}$ ($3000^{\circ}\,\mathrm{s^{-1}}$),
just within this ceiling.  The four rotors are driven from a single-cell
(\SI{3.7}{V} nominal) lithium polymer battery, which constrains motor
selection to high-KV, small-diameter units compatible with the propeller
geometry.

\paragraph{Water-skipping requirements.}
For a successful hop the hydrofoils must, during the brief submersion phase,
generate a net vertical impulse large enough to reverse the vehicle's downward
velocity and return it to free flight.  Blade-element simulation of the
submersion dynamics (described in Chapter~\ref{chapter:dynamic}) shows that
both the hydrofoil inclination angle $\beta$ and the spin rate $\omega_z^*$
at water contact are primary drivers of the ejection velocity; the design
therefore cannot be finalised independently of the flight dynamics.  In
addition, the foils must not generate so much profile drag during flight
that $\omega_z^*$ falls outside the controllable regime of
Chapter~\ref{chapter:papertwo}.

\section{Overall Mechanical Design}
\label{sec:mech_design}

\paragraph{Frame and propulsion.}
The frame is an X-configuration quadrotor with arm length $l =
50\;\mathrm{mm}$ from the centre of mass to each motor mount.  Four
\mbox{0802} motors rated at $22{,}000\;\mathrm{KV}$ are installed at the arm
tips, all rotating in the same direction.  Each motor drives a two-blade
propeller of $50\;\mathrm{mm}$ diameter.  The propeller tip circle therefore
extends to $l + r_p = 50 + 25 = 75\;\mathrm{mm}$ from the vehicle centre.

\paragraph{Electronics.}
The flight controller is a Crazyflie\;2.x Brushless board, which integrates a
six-axis IMU, a radio transceiver, and motor drivers on a single compact
module.  Power is supplied by a $1\mathrm{S}$ ($3.7\;\mathrm{V}$)
lithium polymer cell of capacity $550\;\mathrm{mAh}$, mounted centrally to
minimise the shift of the centre of mass.  The complete vehicle has an all-up
mass of $m = 44\;\mathrm{g}$.

\paragraph{Spin-limiting and hydrofoil assembly.}
At each of the four $45^{\circ}$ azimuthal offsets, a vertical support extends
downward from the frame and carries a two-part element that points radially
inward toward the vehicle centre.  The upper part is a flat \emph{drag plate}
whose face is parallel to the yaw axis, so that as the vehicle spins the plate
sweeps broadside through the air and generates large profile drag; this drag
is the primary means of limiting the passive spin rate
(Section~\ref{sec:design_requirements}).  Directly beneath each drag plate is
a \emph{hydrofoil}, the angled surface that produces hydrodynamic lift during
the brief water-surface contact.  Both elements sit midway between adjacent
propellers so that the downward propeller wash passes between them rather than
directly onto them, avoiding aerodynamic interference during hovering flight.
The hydrofoil tip reaches a radius $r_f = 70\;\mathrm{mm}$ from the vehicle
centre---$20\;\mathrm{mm}$ beyond the motor arm length $l = 50\;\mathrm{mm}$
---placing it just inside the propeller tip circle and beneath the propeller
plane.  The four-element assembly rotates with the vehicle at yaw rate
$\omega_z$, as described in Section~\ref{sec:attitude_dynamics}.

\section{Hydrofoil and Drag-Plate Geometry}
\label{sec:foil_design}

\paragraph{Drag plate.}
The drag plate is a flat rectangular fin mounted vertically---its plane
containing the yaw axis---and extending radially inward from the support.  As
the vehicle spins, the plate presents its broad face to the oncoming air and
acts purely as a spin brake: it generates negligible vertical force but large
yaw-opposing drag, setting the relaxed-hover spin rate $\omega_z^*$ through
the passive balance derived in Section~\ref{sec:attitude_dynamics}.  The plate
is characterised by its planform area $S_d$, its drag coefficient $C_{D,d}$
(close to that of a flat plate in broadside flow), and the effective radius
$r_d$ of its area centroid from the vehicle centre.  These are fixed at
manufacture and chosen so that $\omega_z^*$ falls below the controllability
ceiling $20\pi\;\mathrm{rad\,s^{-1}}$.% TODO: state final S_d, r_d from as-built plates

\paragraph{Hydrofoil profile and planform.}
Each hydrofoil is a thin flat plate, chosen for simplicity of fabrication and
because the flat-plate hydrodynamic model (lift coefficient
$C_L = 2\sin\alpha\cos\alpha$, drag coefficient $C_D = 2\sin^2\alpha$)
provides a closed-form basis for the blade-element simulation of
Chapter~\ref{chapter:dynamic}.  The foil extends radially inward from the tip
radius $r_\mathrm{tip} = r_f = 70\;\mathrm{mm}$ and is parametrised by its
planform area $S_f$ and chord $c_f$ (the dimension tangential to the sweep
direction); for a rectangular foil the radial span is $S_f/c_f$, so the root
radius is
\begin{equation}
    r_\mathrm{root} = r_\mathrm{tip} - \frac{S_f}{c_f}.
    \label{eq:foil_root}
\end{equation}
The fabricated hydrofoil has chord $c_f = 10\;\mathrm{mm}$ and radial span
$30\;\mathrm{mm}$, giving a planform area $S_f = 300\;\mathrm{mm}^2$ and, from
\eqref{eq:foil_root}, a root radius $r_\mathrm{root} = 40\;\mathrm{mm}$.

\paragraph{Attack angle as the design variable.}
The hydrofoil is mounted at a fixed attack angle $\beta$ relative to the
horizontal plane, and this angle is the primary independent variable of the
experimental study.  Candidate values $\beta \in \{5^{\circ}, 10^{\circ},
15^{\circ}, 20^{\circ}, 25^{\circ}, 30^{\circ}\}$ are evaluated against
ejection velocity and hop height, with the blade-element model of
Chapter~\ref{chapter:dynamic} predicting the corresponding water-skipping
performance.  Each value of $\beta$ is realised as a separate fabricated
hydrofoil, and angles are compared by swapping complete foil sets between test
runs.  No active or passive attack-angle mechanism is implemented; $\beta$ is
fixed at manufacture for each foil.

\section{Structural Analysis of the Hydrofoils}
\label{sec:foil_fea}

% TODO: ANSYS FEA of the candidate hydrofoil geometries.
% Loads: peak water-impact distributed pressure extracted from the
%   blade-element model (Chapter~\ref{chapter:dynamic}); root clamped at mount.
% Static structural: max von Mises vs yield (safety factor); tip deflection
%   as fraction of chord (justifies the rigid-foil assumption in the BEM model).
% Modal: first natural frequencies vs spin/impact excitation.
% Geometry + material sweep -> stiffness/strength/mass trade table feeding
%   the foil-set selection in Section~\ref{sec:foil_design}.